---
title: "task_1_decathalon_data"
output: html_notebook
---

```{r}
library(here)
```

```{r}
here::here()
```
# read in data
```{r}
decathalon_data <- read_rds(here("raw_data/decathlon.rds"))
head(decathalon_data)
```
# cleaned headers
```{r}
dec_data_clean_names <- clean_names(decathalon_data)
head(dec_data_clean_names)
```
```{r}
row_names_to_column <- tibble::rownames_to_column(dec_data_clean_names)
```
```{r}
colnames(row_names_to_column)
```
```{r}
rowname_to_names <- row_names_to_column %>% 
  rename(surname = rowname)

rowname_to_names
```

# Q1 Who had the longest long jump seen in the data?
```{r}
dec_data_lc[order(dec_data_lc$long_jump, decreasing = TRUE),]
```
# sorted long jump by decending order, this showed me that clay had the longest jump

# Q2 What was the average 100m time in each competition?
```{r}
dec_data_lc %>% 
  group_by(competition) %>% 
  summarise(avg = mean(x100m))
```
# grouped by competition and calculated the mean of the x100m column

# Q3 Who had the highest total points across both competitions?
```{r}
highest_total_points <- dec_data_lc %>%
  select(surname, points, competition) %>% 
  pivot_wider(names_from = competition,
              values_from = points)
highest_total_points$total_points=rowSums(cbind(highest_total_points$decastar, highest_total_points$olympicg))

head(highest_total_points)
  
               
```
# i took the columns i need and made a new table, i then seperated the two competitions into their own columns, and added the points from each into a totalfor each surname, this gave me the person with the most points.

# Q4 What was the shot-put scores for the top three competitors in each competition?
```{r}
shot_put_scores <- dec_data_lc %>% 
  select(surname, shot_put, competition) %>% 
  pivot_wider(names_from = competition,
              values_from = shot_put) %>% 

```

```{r}
shot_put_scores %>% 
  group_by(olympicg, decastar) %>% 
  slice(1:3)
```



# Q5 What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
